---
title: "Property Retention Model"
author: "Gregory Guranich"
output:
  html_document:
    css: style.css
    toc: true
    toc_depth: 3
    toc_float: true
---
<style type="text/css">
  pre:not([class]) {
    background-color: white;
  }
</style>


<style type="text/css">
h1 {
  font-size: 34px;
}
h1.title {
  font-size: 60px;
}
h2 {
  font-size: 22px;
}
h3 {
  font-size: 16px;
}
h4 {
  font-size: 14px;

h6 {
  font-size: 12px;
}
.table th:not([align]) {
  text-align: left;
}
</style>

#### 1) Business problem

#### You work for Kangaroo Auto Insurance Company, an Australian company. Your business partner, who is not at all familiar with statistics, would like you to create a retention model based on historical policy data. Your business partner is concerned about the company’s retention rates as well as the key drivers that cause policy cancellations.
#### For this case competition, your group is tasked with identifying policies that will cancel before the end of their term. With the help of a predictive model, your group will be able to predict those policies that are most likely to cancel as well as understand what variables are most influential in causing a policy cancellation.

```{r setup, include=FALSE}

library('ggplot2') # plots... not used at the moment
library('polynom') # what is this for?
library('plyr') # data scrub
#library(rjags)
#library(R2jags) # bayes model not seen in this file
library('glmnet') # for glm model
library('dplyr') #for data scrub
library("caret") #for dummy cariables
#library(VIM) # imputation, not used here
#library(mice) # na plot
library(corrplot) #cor plot
library('pROC') #roc calc
library(ROCR) # for bagging
library(foreach) # for bagging
library('e1071') # for svm
#-------------------
Logit <- function(p) log(p/(1-p))
InvLogit <- function(x) 1/(1+exp(-x))

DerivativeInvLogit <- function(x) 1/(1+exp(-x))^2*exp(-x)
DerivativeInvLogit <- function(x, beta0=0, beta1 =1) 1/(1+exp(-(beta0+beta1*x)))^2*beta1*exp(-(beta0+beta1*x))



# Import Data, Merge Data
Validation <- read.csv("test.csv")
states <- read.csv("free-zipcode-database-Primary.csv")
MergedTestData <- merge(Validation,states, by.x = "zip.code", by.y = "Zipcode", all.x = TRUE)
trav_test <- MergedTestData

# To train model
trav_orig <- read.csv('train.csv') # keep original set for NA analysis
MergedTestData <- merge(trav_orig,states, by.x = "zip.code", by.y = "Zipcode", all.x = TRUE)
trav_complete <- MergedTestData

```
## Data Cleaning
### NA Plots
```{r}
# NA imputation, DO THIS BEFORE OMIT.NA
mice_plot <- aggr(trav_orig, col=c('navyblue','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(trav_orig), cex.axis=.7,
                    gap=1, ylab=c("Missing data","Pattern"))

```


### Imputation
```{r}
# As you saw in the plot above the NAs were created in a pattern (artificial data)
# This is extra reason to impute the NAs
# Because of the pattern we no overlap among NAs
# This allows us to use other correlated variables to predict what the NAs might be
# Unfortunately for this part, very few of the variables are correlated

# Sales.Channel is correlated with our repsonse cancelation
# Here we see NA's have about the same rate of cancelation as, phone, and online, 
# we will impute these NAs by binning all of the above groups (this is sometiems bad practice, consider changing this later)
sum(is.na(trav_complete$sales.channel))

channel<-ifelse(trav_complete$sales.channel=="Broker", 1, 0)
channel[(is.na(channel))]<-0
#sum(is.na(channel)) #just to check
trav_complete$channel<-channel
```
### Erroneous data
```{r}
# fixing erroneous cancel entry of -1
summary(trav_complete$cancel)
neg <- trav_complete$cancel==-1
trav_complete$cancel[neg] <- NA
# fixing erroneous age entry
# Lets assume age was typed in wrong if over 117 for example 159 could be 59 or 225 could be 25
# chose 117 because someone who is 18 might have retners insurance if 118=18 this is room for possible typo
a<-trav_complete$ni.age>117
trav_complete$ni.age[a]<-NA

## Removal
# We need to remove the merged data first because variables such as City had 1000s of NAs
names(trav_complete)
# OK now we can finally omit the remaining NAs, amros merged set has a bunch of nas so we need to remove all those columns first


trav_complete<-trav_complete[,c("cancel","credit","channel","claim.ind","ni.age","tenure","State","year","ni.marital.status",
                                "n.children","len.at.res","n.adults","dwelling.type","house.color","coverage.type",'premium')]
#"premium","City","len.at.res","dwelling.type","n.adults","EstimatedPopulation" for use later
sum(is.na(trav_complete))
colnames(trav_complete)[7]<-"state"
trav_complete<-na.omit(trav_complete)
```
## Variable Selection
```{r}
#The end goal is to dummy encode in order to select reference categories
#We must start by making data categorical
#We will use out-of-fold likelihood method
#Making continuous variables discrete in order to reveal trends
```

### Adults
```{r}
table(trav_complete$cancel,trav_complete$n.adults)
tab<-table(trav_complete$cancel,trav_complete$n.adults)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
# cancellation rate for each category
rate

# Lets bin 5+ adults to make a decent size bin
trav_complete <- trav_complete %>% mutate(adults=cut(n.adults, breaks=c(1,2,3,4,5, Inf), labels=c(1,2,3,4,5)))

#### lets make a general table / rate code
tab<-table(trav_complete$cancel,trav_complete$adults)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
# cancellation rate for each category
rate
x<-c(1,2,3,4,5)
plot(x,rate, pch=16, xlab="adults", ylab="cancelation %")
grid()
lines(rate~x, lwd=3, col='red')
```


```{r}
# Rates are similar in first 3 bins, lets simplify categories here
trav_complete <- trav_complete %>% mutate(adults=cut(n.adults, breaks=c(-Inf,3,4,5, Inf), labels=c(1,2,3,4)))

table(trav_complete$cancel,trav_complete$adults)
tab<-table(trav_complete$cancel,trav_complete$adults)
rate<-(tab[2,]/tab[1,])
# cancellation rate for each category
rate
```
### Length of Residence
```{r}

# lets break it to descrete 1-30
trav_complete <- trav_complete %>% mutate(length=cut(len.at.res, breaks=c(-Inf, seq(1,30,1), Inf), labels=c(seq(1,31,1))))
table(trav_complete$cancel,trav_complete$length)
tab<-table(trav_complete$cancel,trav_complete$length)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
rate


rate
x<-seq(1,31,1)
plot(x,rate, pch=16, xlab="Length of Residence", ylab="cancelation %")
grid()
lines(rate~x, lwd=3, col='red')
# we see a drop at len10
# we see a drop at len19
# goes up at 22
# sample size is too low in these bins to see whats going on

# lets group the tail ends to get a different look
trav_complete <- trav_complete %>% mutate(length=cut(len.at.res, breaks=c(-Inf, seq(7,21,1), Inf), labels=c(seq(1,16,1))))
table(trav_complete$cancel,trav_complete$length)
tab<-table(trav_complete$cancel,trav_complete$length)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
x<-seq(1,16,1)
plot(x,rate, pch=16, xlab="Length of Residence", ylab="cancelation %")
grid()
lines(rate~x, lwd=3, col='red')

# lets make the strong groups, we dont have enough evidence about this upper tail
# first break is more important
# trav_complete <- trav_complete %>% mutate(length=cut(len.at.res, breaks=c(-Inf,10,15,20,Inf), labels=c(seq(1,4,1))))
trav_complete <- trav_complete %>% mutate(length=cut(len.at.res, breaks=c(-Inf,10,18,Inf), labels=c(seq(1,3,1))))
table(trav_complete$cancel,trav_complete$length)
tab<-table(trav_complete$cancel,trav_complete$length)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
x<-seq(1,3,1)
plot(x,rate, pch=16, xlab="Length of Residence", ylab="cancelation %")
grid()
lines(rate~x, lwd=3, col='red')


plot(density(trav_complete$len.at.res), main="Length at Residence, pdf")
abline(v=10, lty=2)
abline(v=18, lty=2)

```
### Dwelling Type
```{r}


table(Validation$credit, Validation$dwelling.type)

# We see that these landlords are not new customers it is just a new product for homeowners
# Most of them have high credit so they probably behave similar to homeowners
# this variable probably wont be very helpful

# Dwelling type ~ tenant
tenant<-ifelse(trav_complete$dwelling.type=="Tenant", 1, 0)
trav_complete$tenant<-tenant

# Our intuition took us here to look for a relationship between dwelling and length at residence
# our intuition says length at residence wont matter if they are a renter
trav_complete <- trav_complete %>% mutate(length2=cut(len.at.res, breaks=c(-Inf, seq(7,21,1), Inf), labels=c(seq(1,16,1))))

response.i <- as.numeric(as.character((trav_complete$cancel)))
mar.i <- as.numeric(trav_complete$tenant)
kids.i <- as.numeric(trav_complete$length2)
summary(kids.i)
# exploratory plot
res0.t <- tapply(response.i[mar.i==0], kids.i[mar.i==0], mean)
res1.t <- tapply(response.i[mar.i==1], kids.i[mar.i==1], mean)
par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,1))
plot(res0.t ~ c(seq(1,16,1)), type = "l", ylim = c(0, max(res0.t, res1.t)), ylab = "% cancel", xlab = "# length at res", xlim=c(0,20))
lines(res1.t ~ c(seq(1,16,1)), col = 2, lty = 2)
legend("topright", legend = c("tenant==0", "tenant==1"), col = c(1,2), lty = c(1,2), cex = 1.5)
```

### Children
```{r}
# Our intuition took us here to look for a relationship

# We are also suspicious about the 0 correlation between kids and marriage when we expected high corr

response.i <- as.numeric(trav_complete$cancel)
mar.i <- as.numeric(trav_complete$ni.marital.status)
kids.i <- as.numeric(trav_complete$n.children)
summary(kids.i)
# exploratory plot
res0.t <- tapply(response.i[mar.i==0], kids.i[mar.i==0], mean)
res1.t <- tapply(response.i[mar.i==1], kids.i[mar.i==1], mean)
par(lwd = 3, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5,mar = c(5,5,1,1), mfrow = c(1,1))
plot(res0.t ~ c(0,1,2), type = "l", ylim = c(0, max(res0.t, res1.t)), ylab = "% cancel", xlab = "# kids", xlim=c(0,12))
lines(res1.t ~ c(0,1,2,3,4,5,6,7,8,9,10,11,12), col = 2, lty = 2)
legend("topright", legend = c("marriage==0", "marriage==1"), col = c(1,2), lty = c(1,2), cex = 1.5)


### Children
trav_complete <- trav_complete %>% mutate(children=cut(n.children, breaks=c(-Inf,2,5,9,Inf), labels=c(1,2,3,4)))

```
### Age

```{r}

# Lets make some typical groups to look at 

trav_complete <- trav_complete %>% mutate(age_g=cut(ni.age, breaks=c(-Inf, 25, 35, 45, 55, 65, 75, Inf), labels=c(1,2,3,4,5,6,7)))
table(trav_complete$cancel,trav_complete$age_g)
tab<-table(trav_complete$cancel,trav_complete$age_g)
rate<-(tab[2,]/(tab[1,]+tab[2,]))

x<-seq(1,7,1)
plot(x,rate, pch=16, xlab="Age", ylab="cancelation %")
grid()
lines(rate~x, lwd=3, col='red')

# lets look at the interaction between the broker and age
one=trav_complete$channel==1
tab<-table(trav_complete$cancel[one],trav_complete$age_g[one])
rate<-(tab[2,]/(tab[1,]+tab[2,]))
tab<-table(trav_complete$cancel[!one],trav_complete$age_g[!one])
rate2<-(tab[2,]/(tab[1,]+tab[2,]))

plot(x, ylim=c(.1,.4), pch=16, xlab="Age", ylab="cancelation %", type="n")
grid()
lines(rate~x, lwd=3, col='red')
lines(rate2~x, lwd=3, col="blue")
# there appears to be a strong interaction here
```





```{r}
trav_complete <- trav_complete %>% mutate(age_g=cut(ni.age, breaks=c(-Inf, 35, 55, Inf), labels=c(1,2,3)))
table(trav_complete$cancel,trav_complete$age_g)
tab<-table(trav_complete$cancel,trav_complete$age_g)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
rate
# Lets adjust these groups a bit so that the jump in cancelation rate is even between breaks
# These steps make the new variable ordinal which is better for correlation detection and model fitting

trav_complete <- trav_complete %>% mutate(age_g=cut(ni.age, breaks=c(-Inf, 30, 55, Inf), labels=c(1,2,3)))
table(trav_complete$cancel,trav_complete$age_g)
tab<-table(trav_complete$cancel,trav_complete$age_g)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
rate
```
### Tenure
```{r}
plot(density(na.omit(trav_orig$tenure)), main="Tenure Bimode")
# This is bi modal so we can make a dichotomous variable
trav_complete <- trav_complete %>% mutate(tenure_g=cut(tenure, breaks=c(-Inf, 9, Inf), labels=c(0,1)))
table(trav_complete$cancel,trav_complete$tenure_g)
tab<-table(trav_complete$cancel,trav_complete$tenure_g)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
rate
# I hope this works this makes a nice even break ...

# Lets still look at some tables
table(trav_complete$cancel,trav_complete$tenure)
tab<-table(trav_complete$cancel,trav_complete$tenure)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
rate
# I see a sweet spot for tenure 9, 10, 11 ... not intuitive
trav_complete <- trav_complete %>% mutate(tenure_g=cut(tenure, breaks=c(-Inf,8, 11, Inf), labels=c(1,2,3)))
table(trav_complete$cancel,trav_complete$tenure_g)
tab<-table(trav_complete$cancel,trav_complete$tenure_g)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
rate
# Very nice breaks! but the order is off
# Can we just change the order... will that work as ordinal?
trav_complete <- trav_complete %>% mutate(tenure_g=cut(tenure, breaks=c(-Inf,8, 11, Inf), labels=c(1,3,2)))
plot(density(na.omit(trav_orig$tenure)), main="Tenure Phenom")
abline(v=8, lty=2)
abline(v=11, lty=2)


# Here is a very complicated version with the danger zone
trav_complete <- trav_complete %>% mutate(tenure_g=cut(tenure, breaks=c(-Inf,1,3,8, 11, Inf), labels=c(1,2,3,4,5)))
table(trav_complete$cancel,trav_complete$tenure_g)
tab<-table(trav_complete$cancel,trav_complete$tenure_g)
rate<-(tab[2,]/(tab[1,]+tab[2,]))
rate
plot(density(na.omit(trav_orig$tenure)), main="Tenure Phenom")
abline(v=1, lty=2, col=2)
abline(v=3, lty=2, col=2)
abline(v=8, lty=2, col=3)
abline(v=11, lty=2, col=3)

tenure_ord<-ifelse(trav_complete$tenure_g==1|trav_complete$tenure_g==3|trav_complete$tenure_g==5, 2 ,0)
safe<-trav_complete$tenure_g==4
unsafe<-trav_complete$tenure_g==2
tenure_ord[safe]<-1
tenure_ord[unsafe]<-3
trav_complete$tenure_ord<-tenure_ord
```
### state
```{r}
trav_complete$state<-as.character(trav_complete$state)
table(trav_complete$state)
premtab<-(table(trav_complete$cancel, trav_complete$state))
rate<-(premtab[2,]/(premtab[1,]+premtab[2,]))
rate


trav_complete$state2 <- trav_complete$state
#Lets combine states using out of fold likelihood
#a<-trav_complete$state2=="VA"
#b<-trav_complete$state2=="WA"
#c<-trav_complete$state2=="AZ"
#trav_complete$state2[a|b|c] <- "VA/WA/AZ"


b<-trav_complete$state2=="WA"
c<-trav_complete$state2=="AZ"
trav_complete$state2[b|c] <- "WA/AZ"



a<-trav_complete$state2=="IA"
b<-trav_complete$state2=="CO"
trav_complete$state2[a|b] <- "IA/CO"

premtab<-(table(trav_complete$cancel, trav_complete$state2))
rate<-(premtab[2,]/(premtab[1,]+premtab[2,]))
rate

trav_complete$state2 <- as.factor(trav_complete$state2)
trav_complete$state <- as.factor(trav_complete$state)

sum(is.na(trav_complete$state))
```



### Premium
```{r}
# Lets break premium into
trav_complete <- trav_complete %>% mutate(premium_g=cut(premium, breaks=c(-Inf,826,852,875,896,920,966,995, Inf), labels=c(seq(1,8,1))))
table(trav_complete$cancel, trav_complete$premium_g)
premtab<-(table(trav_complete$cancel, trav_complete$premium_g))
rate<-(premtab[2,]/premtab[1,])
# cancellation rate for each category

rate
# 896,920 prem greatest jump in cancelation
# It seems there is a range of premiums in which cancelation rates are low
# Maybe these are customers being priced correctly?

# lets create simplified levels
trav_complete <- trav_complete %>% mutate(premium_g=cut(premium, breaks=c(-Inf,852,875,896,920,966,Inf), labels=c(seq(1,6,1))))
table(trav_complete$cancel, trav_complete$premium_g)
premtab<-(table(trav_complete$cancel, trav_complete$premium_g))
rate<-(premtab[2,]/premtab[1,])
# cancellation rate for each category
rate



trav_complete <- trav_complete %>% mutate(premium_g=cut(premium, breaks=c(-Inf,852,966,Inf), labels=c(seq(1,3,1))))
premtab<-(table(trav_complete$cancel, trav_complete$premium_g))
rate<-(premtab[2,]/premtab[1,])
rate
out<-ifelse(trav_complete$premium_g==2,1,0)
trav_complete$premium_g<-out

```


## LASSO
```{r}
# Elastic Net
## LASSO

# We can only use categories here
names(trav_complete)
# removed house color, coverage.type

cats<-(trav_complete[,c("credit","ni.marital.status","children","channel","age_g")])


cats<-(trav_complete[,c("credit","ni.marital.status","children","channel","claim.ind","age_g","tenure_ord", "state2","year","adults","length","tenant","premium_g")])
cats<-sapply(cats, factor)
str(cats)

cancel<-trav_complete$cancel
# We need dummy variables for an elastic net
dummy<-dummyVars(~.,data=cats, fullRank = TRUE)
dummyset<-as.data.frame(predict(dummy, cats))
#write.csv(trav2, "full_rank_dummy.csv")
#dummyset$cancel<-cancel

# Make your train and test vectors, data must be a matrix
train.x<-data.matrix(dummyset)
train.y<-trav_complete[,c("cancel")]

# Lets set elastic net (alpha) to 1, this is LASSO regression
CV=cv.glmnet(x=train.x ,y=train.y, family='binomial', type.measure = 'auc', alpha=1, nlambda=100)

plot(CV)

# lets take the minimum error, most regularlized 1se down from min error
fit=glmnet(x=train.x,y=train.y, family='binomial', alpha=1, lambda=CV$lambda.1se)
nonzero<-fit$beta[,1]>0.001
beta<-fit$beta[,1]
beta<-beta[nonzero]
beta[order(beta)]
fit

library(gglasso)

group=c(1,1,2,3,3,3,4,5,5)
train.x<-as.matrix(dummyset)
train.y2<-ifelse(train.y==0,-1,1)


cv <- cv.gglasso(x=train.x, y=train.y2, group=group, loss="logit",
pred.loss="L2", lambda.factor=0.05,nfolds=10)


fit <- gglasso(x=train.x, y=train.y2, group=group, loss="logit")



coef.mat=fit$beta

#Group1 enters the equation
g1=max(which(coef.mat[1,]==0))

#Group2 enters the equation
g2=max(which(coef.mat[3,]==0))

#Group3 enters the equation
g3=max(which(coef.mat[4,]==0))

#Group4 enters the equation
g4=max(which(coef.mat[7,]==0))

g5=max(which(coef.mat[8,]==0))


#Coefficient Plot. Let's also use some nice colors
library(RColorBrewer)
cols=brewer.pal(5,name="Set1")

plot(fit$b0,main="Coefficient vs Step",
     ylab="Intercept",xlab="Step (decreasing Lambda =>)",
     col=cols[1],
     xlim=c(-1,100),
     ylim=c(.5,-1.5),
     type="l",lwd=4)
grid()
par(new=T)

x=c(g1,g2,g3,g4,g5)
y=c(fit$b0[g1],fit$b0[g2],fit$b0[g3],fit$b0[g4],fit$b0[g5])

plot(x=x,y=y,pch=13,lwd=2,cex=2,col=cols[-1],
     xlim=c(-1,100),ylim=c(.5,-1.5),
     xaxt='n',yaxt='n',xlab="",ylab="")

lmda=round(fit$lambda[c(g1,g2,g3,g4,g5)],3)
text(x=x-0.05,y=y+0.01,labels=c("Group1","Group2","Group3","Group4","Group5"),pos=3,cex=0.9)
text(x=x-0.05,y=y-0.01,labels=paste("Lambda\n=",lmda),pos=1,cex=0.8)



lmbda=cv$lambda.1se
(coefs=coef.gglasso(object=fit,s=lmbda))
plot(cv)


predict.gglasso(object=fit,newx=X,s=lmbda)

```


```{r}

library(glinternet)


cats<-(trav_complete[,c("credit","ni.marital.status","children","channel","claim.ind","age_g","tenure_ord", "state2","year","adults","length","tenant","premium_g")])

# hand pick columns to drop with the largest sample size
table(cats$credit)#high
table(cats$ni.marital.status)#1
table(cats$children)#1
table(cats$channel)#1
table(cats$claim.ind)#0
table(cats$age_g)#2
table(cats$tenure_ord)#2
table(cats$state2)#IA/CO
table(cats$year)#2016
table(cats$adults)#1
table(cats$length)#2
table(cats$tenant)#0
table(cats$premium_g)#0




cats<-sapply(cats, factor)

dummy<-dummyVars(~.,data=cats, fullRank = FALSE)
dummyset<-as.data.frame(predict(dummy, cats))



# remove columns hand picked for reference categories
dummyset <- dummyset[,-c(1,5,6,11,12,15,18,20,27,28,33,35,37)]

#now 25 vars/ columns



```



```{r}
cats$credit<-factor(as.numeric(cats$credit))
levels(cats$credit)<-c(0,1,2)
levels(cats$children)<-c(0,1,2,3)
levels(cats$age_g)<-c(0,1,2)

cats<-sapply(cats, factor)
train.x<-data.matrix(cats)
train.y<-trav_complete[,c("cancel")]


group=c(3,2,4,2,3)

fit <- glinternet(X=train.x, Y=train.y, numLevels=group, nFolds = 5, lambda=NULL, nLambda=50,
lambdaMinRatio=0.01, interactionCandidates=NULL, family="binomial")

cv <- glinternet.cv(X=train.x, Y=train.y, numLevels=group, nFolds = 5, lambda=NULL, nLambda=50,
lambdaMinRatio=0.01, interactionCandidates=NULL, family="binomial")


fit$betahat
coef(object, lambdaIndex = NULL, ...)

lambda <- cv$lambda
which(cv$lambda %in% cv$lambdaHat1Std)
lambda.index <- which(cv$lambda %in% cv$lambdaHat1Std)


plot(cv)

coef(object=cv, lambdaIndex = lambda.index)

14











```
```{r}
# Coverage type isnt doing enough it could overfit the model
# length makes sense and has a significant impact
# adult group 4 has an impact, maybe we can make this dichotomous
# see model below for p values
modelset2 <- dummyset
model <- glm(cancel.1~.,data=modelset2, family = "binomial"(link="logit"))
summary(model)
```
#### In conclusion age can be dichotomous. It only seems to matter if they are older than 55. My intuition is that the younger age is already captured by lower credit. Adults can be lowered to 3 categories. Lasso tells us we will not recieve lift between the first two categories. Lets consider using the tails as danger zones. Coverage and house color will not provide lift. Lets focus on lift for this part of the competition but we may regress to a lower performance model in order to have an explanatory model.
#### reduction; age, adults, tenure, state IA

## ANOVA
```{r}
# For this stage we need to use our intuition to order variables in the model
# The order should be nested or structurally dependent
# With the first variable being the top of the hierachy

model <- glm(cancel~credit+channel+family_level+adults+age_g+tenure_ord+length+state+claim.ind+year+premium_g ,data=cats, family = "binomial"(link="logit"))
summary(model)
anova(model, test="Chisq")
```

## Validation
### Before Tweeks
```{r}
cats <- as.data.frame(cats)
cats2<-(cats[,c('cancel','credit','channel','family_level','adults','age_g','tenure_ord','state','claim.ind','year','premium_g','length')])

dummy<-dummyVars(~.,data=cats2, fullRank = TRUE)
dummyset<-as.data.frame(predict(dummy, cats2))

df<-dummyset
data.size<-nrow(df)
train.size<-0.8
train.row.nums<-sample(1:data.size, data.size*train.size, replace=FALSE)
test.row.nums<-setdiff(1:data.size,train.row.nums)

train.data<-subset(df[train.row.nums,])
test.data<-subset(df[test.row.nums,])

model <- glm(cancel.1~., data=train.data, family = "binomial"(link="logit"))
#model_save <- glm(cancel.1~.,data=modelset2, family = "binomial"(link="logit"))

test<-test.data[,c(1)]

predictions<-predict(objec = model, newdata = test.data[,-c(1)], type='response')

auc <- roc(test, predictions)
auc

pred<-prediction(predictions, test.data[,c(1)])
perf<-performance(pred,'tpr','fpr')
plot(perf,main="ROC")
abline(0,1,lty=3)
```


```{r, include=FALSE}
# Import Data, Merge Data
Validation <- read.csv("test.csv")
states <- read.csv("free-zipcode-database-Primary.csv")
MergedTestData <- merge(Validation,states, by.x = "zip.code", by.y = "Zipcode", all.x = TRUE)
trav_test <- MergedTestData

# To train model
trav_orig <- read.csv('train.csv') # keep original set for NA analysis
MergedTestData <- merge(trav_orig,states, by.x = "zip.code", by.y = "Zipcode", all.x = TRUE)
#MergedTestData2 <- merge(MergedTestData, Pop, by.x = "zip.code", by.y = "zip.code", all.x = TRUE)


trav_complete <- MergedTestData


sum(is.na(trav_complete$sales.channel))

channel<-ifelse(trav_complete$sales.channel=="Broker", 1, 0)
channel[(is.na(channel))]<-0
#sum(is.na(channel)) #just to check
trav_complete$channel<-channel

## Erroneous and NA
# fixing erroneous cancel entry of -1
summary(trav_complete$cancel)
neg <- trav_complete$cancel==-1
trav_complete$cancel[neg] <- NA
# fixing erroneous age entry
# Lets assume age was typed in wrong if over 117 for example 159 could be 59 or 225 could be 25
# chose 117 because someone who is 18 might have retners insurance if 118=18 this is room for possible typo
a<-trav_complete$ni.age>117
trav_complete$ni.age[a]<-NA



## Removal
# We need to remove the merged data first because variables such as City had 1000s of NAs
names(trav_complete)
#a<-is.na(trav_complete$pop)
#trav_complete$pop[a]<-0
# OK now we can finally omit the remaining NAs, amros merged set has a bunch of nas so we need to remove all those columns first
trav_complete<-trav_complete[,c("cancel","credit","channel","claim.ind","ni.age","tenure","state","year","ni.marital.status",
                                "n.children","len.at.res","n.adults",'premium')]
#"premium","City","len.at.res","dwelling.type","n.adults","EstimatedPopulation" for use later
sum(is.na(trav_complete))
trav_complete<-na.omit(trav_complete)

## Variable Selection


### Premium
trav_complete <- trav_complete %>% mutate(premium_g=cut(premium, breaks=c(-Inf,852,966,Inf), labels=c(seq(1,3,1))))
premtab<-(table(trav_complete$cancel, trav_complete$premium_g))
rate<-(premtab[2,]/premtab[1,])
rate
out<-ifelse(trav_complete$premium_g==2,1,0)
trav_complete$premium_g<-out


### Adults
trav_complete <- trav_complete %>% mutate(adults=cut(n.adults, breaks=c(-Inf, 5, Inf), labels=c(0,1)))
premtab<-(table(trav_complete$cancel, trav_complete$adults))
rate<-(premtab[2,]/premtab[1,])
rate


### Length of Residence
trav_complete <- trav_complete %>% mutate(length=cut(len.at.res, breaks=c(-Inf,10,18,Inf), labels=c(seq(1,3,1))))

### Children
trav_complete <- trav_complete %>% mutate(children=cut(n.children, breaks=c(-Inf,2,5,9,Inf), labels=c(1,2,3,4)))

### Age
trav_complete <- trav_complete %>% mutate(age_g=cut(ni.age, breaks=c(-Inf, 55, Inf), labels=c(0,1)))
#trav_complete <- trav_complete %>% mutate(age_g=cut(ni.age, breaks=c(-Inf, 30, 55, Inf), labels=c(1,2,3)))


### Tenure
a<-trav_complete$year=="2013"
b<-trav_complete$year=="2014"
c<-trav_complete$year=="2015"
d<-trav_complete$year=="2016"

plot(density(trav_complete$tenure[a]))
plot(density(trav_complete$tenure[b]))
plot(density(trav_complete$tenure[c]))
plot(density(trav_complete$tenure[d]))

hist(trav_complete$tenure[a])
hist(trav_complete$tenure[b])
hist(trav_complete$tenure[c])
hist(trav_complete$tenure[d])

#can we put a tracker on these people through a new variables
#track<-trav_complete$year=="2013"&trav_complete$tenure==c(8,9,10)
```


```{r}
trav_complete <- trav_complete %>% mutate(tenure_g=cut(tenure, breaks=c(-Inf,1,3,8, 11, Inf), labels=c(1,2,3,4,5)))
table(trav_complete$cancel,trav_complete$tenure_g)

tenure_ord<-ifelse(trav_complete$tenure_g==1|trav_complete$tenure_g==3|trav_complete$tenure_g==5, 2 ,0)
safe<-trav_complete$tenure_g==4
unsafe<-trav_complete$tenure_g==2
tenure_ord[safe]<-1
tenure_ord[unsafe]<-3
trav_complete$tenure_ord<-tenure_ord



### state
table(trav_complete$state)
trav_complete$state<-as.character(trav_complete$state)
table(trav_complete$state, trav_complete$cancel)

tab<-table(trav_complete$cancel, trav_complete$state)
rate<-(tab[2,]/tab[1,])
rate
#VA and WA are similar in location, lets make a larger bin with them
a<-trav_complete$state=="VA"
b<-trav_complete$state=="WA"
trav_complete$state[a|b] <- "VA/WA"
#a<-trav_complete$state=="IA"
#trav_complete$state[a] <- "AZ"

trav_complete$state <- as.factor(trav_complete$state)
#write.csv(trav_complete, "explore.csv")
# use channel and premium and familylevel
sum(is.na(trav_complete$state))

### YEAR
x<-as.factor(trav_complete$year)
x<-revalue(x, c("2013"="1", "2016"="1"))
x<-revalue(x, c("2014"="3", "2015"="2"))
trav_complete$year<-x

# We can only use categories here
names(trav_complete)

### POpulation type

#trav_complete <- trav_complete %>% mutate(pop_type=cut(pop, breaks=c(-Inf,50000,Inf), labels=c(1,2)))
#tab<-table(trav_complete$cancel, trav_complete$pop_type)
#rate<-(tab[2,]/tab[1,])
#rate

cats<-(trav_complete[,c("cancel","credit","family_level","channel","claim.ind","age_g","tenure_ord","state","year",
                        "length","premium_g","adults")])
cats<-lapply(cats, factor)
str(cats)
# We need dummy variables for an elastic net
dummy<-dummyVars(~.,data=cats, fullRank = TRUE)
dummyset<-as.data.frame(predict(dummy, cats))
#write.csv(trav2, "full_rank_dummy.csv")





# Make your train and test vectors, data must be a matrix
train.x<-data.matrix(dummyset[,-c(1)])
train.y<-trav_complete[,c("cancel")] 

# Lets set elastic net (alpha) to 1, this is LASSO regression
CV=cv.glmnet(x=train.x ,y=train.y, family='binomial', type.measure = 'class', alpha=1, nlambda=100)

plot(CV)

# lets take the minimum error, most regularlized 1se down from min error
fit=glmnet(x=train.x,y=train.y, family='binomial', alpha=1, lambda=CV$lambda.1se)
fit$beta[,1]
fit

```

### Validation Revised GLM
```{r}
model <- glm(cancel~credit+channel+family_level+adults+age_g+tenure_ord+length+state+claim.ind+year+premium_g+channel*age_g ,data=cats, family = "binomial"(link="logit"))
summary(model)
anova(model, test="Chisq")

cats<-as.data.frame(cats)

df<-cats
data.size<-nrow(df)
train.size<-0.8
train.row.nums<-sample(1:data.size, data.size*train.size, replace=FALSE)
test.row.nums<-setdiff(1:data.size,train.row.nums)

train.data<-subset(df[train.row.nums,])
test.data<-subset(df[test.row.nums,])
test<-test.data[,c(1)]


model <- glm(cancel~credit+channel+family_level+adults+age_g+tenure_ord+length+state+claim.ind+year+premium_g+channel*age_g, data=train.data, family = "binomial"(link="logit"))





predictions<-predict(object = model, newdata = test.data[,-c(1)], type='response')
# Plot model
#library('pROC')

auc <- roc(test, predictions)
auc
```
### Bagged GLM
```{r}

train<-train.data
test<-test.data
f<-cancel~credit+channel+family_level+adults+age_g+tenure_ord+length+state+claim.ind+year+premium_g+channel*age_g
length_divisor<-7  
iterations<-15 

predictions<-foreach(m=1:iterations,.combine=cbind) %do% {  
  training_positions <- sample(nrow(train), size=floor((nrow(train)/length_divisor)))  
  train_pos<-1:nrow(train) %in% training_positions
  fit_bag<-glm(f, data=train, family=binomial(logit))
  predict(fit_bag, newdata=test[,-c(1)], type="response")  
}  

p <- rowMeans(predictions)
auc <- roc(test[,c(1)], p)
auc


pred<-prediction(p,test[,c(1)])
perf<-performance(pred,'tpr','fpr')

plot(perf,main="ROC")
abline(0,1,lty=3)
```

### GBM
```{r, include=FALSE, echo=TRUE}
# Uses same train/test as above
train.data<-subset(df[train.row.nums,])
test.data<-subset(df[test.row.nums,])

x<-train.data[,-c(1)]
y<-as.factor(train.data[,c(1)])

cancel2<- as.factor(ifelse(y==0,'no','yes'))
y<-cancel2

levels(y)



model_test <- train(x,
  y,
  method = "gbm",
  metric = "ROC",
  trControl = trainControl(method = "cv", number = 4, summaryFunction =  twoClassSummary, classProbs = TRUE),
  preProcess = c("center", "scale")
)
```
```{r}
summary(model_test)

predictions<-predict(object = model_test, test.data[,-c(1)], type='prob')
# Plot model
plot(model_test)

auc <- roc(test.data[,c(1)], predictions[[2]])

print(auc$auc)
```

```{r, eval=FALSE, include=FALSE}

# For final predictions
tune_out<-tune(svm, cancel~credit+channel+family_level+adults+age_g+tenure_ord+length+state+claim.ind+year+premium_g+channel*age_g,
  data=train.data,
  kernel="radial",
   ranges = list(epsilon = seq(0,1,0.1), cost = 2^(2:9)),
  probability=T
  )
print(tune_out)

tune_out<-tune(svm, cancel~credit+channel+family_level+adults+age_g+tenure_ord+length+state+claim.ind+year+premium_g+channel*age_g,
  data=train.data,
  kernel="radial",
   ranges = list(epsilon = seq(0,.1,.01) cost = 2^(2:9)),
  probability=T
  )
print(tune_out)

bestmodel=tune_out$best.model
summary(bestmodel)


p<-predict(bestmodel, test.data[,-c(1)], type="class", probability=T)
probs_valid<-attr(p, 'probabilities')
probs_valid
predictions_test <- probs_valid[,1]
auc <- roc(test, predictions_test)
auc
plot(tune_out)
# .671! thats terrible
# do I have the kernal wrong? should it be SVR?
```